---
title: "Practical Implementation of the Kelly Criterion"
subtitle: "Reproduction and Statistical Reassessment of Kelly Literature"
author:
  - Syed Bashir Hydari (et al.)
date: today
format:
  pdf:
    toc: true
    toc-depth: 2
    number-sections: true
    fig-width: 7
    fig-height: 2.8
    fig-pos: "H"
    geometry:
      - margin=0.95in
    include-in-header:
      text: |
        \usepackage{float}
        \usepackage{booktabs}
        \usepackage{caption}
        \captionsetup[figure]{font=small,skip=2pt}
        \captionsetup[table]{font=small,skip=2pt}
        \setlength{\parskip}{4pt}
        \setlength{\abovedisplayskip}{3pt}
        \setlength{\belowdisplayskip}{3pt}
execute:
  warning: false
  message: false
  echo: false
  cache: false
editor: source
---

\begin{center}
\textbf{Role Contributions} \\[0.5em]
\begin{tabular}{rl}
Syed Bashir Hydari: & core reproduction, source code, final paper, speaker \\
Aniqa Nayim: & core reproduction, source code, final paper, speaker \\
A \\
B \\
C \\
D \\
E \\
F \\
G \\
H \\
\end{tabular}
\end{center}
\newpage

# Introduction

The Kelly Criterion (Kelly 1956) prescribes the growth-optimal allocation $f^* = \Sigma^{-1}(\mu - r_f \mathbf{1})$ but is sensitive to estimation noise, the mean-vs-median divergence of long-hori[...]

This study has four objectives. *(O1)* Validate Kelly geometry under known GBM parameters via Monte Carlo across four horizons. *(O2)* Replicate CC20's pipeline — single-equity Banca Intesa, in-[...]

The central finding is that CC20's ranking is not robust. The long-only budget cap mechanically collapses Full, Double, and Triple Kelly onto an *identical* portfolio, so the apparent dominance of[...]

# Data and Methodology

We use 35 surviving EuroStoxx 50 constituents with daily adjusted closes from Yahoo Finance (`tidyquant`, `quantmod`), January 2000 through December 2018, after a 20%-missing filter, forward-fill,[...]

The multi-asset Kelly QP follows CC20's Taylor expansion of $\mathbb{E}[\log(1+\mathbf{w}^\top\mathbf{r})]$:
$$
\max_{\mathbf{w}} \; \mathbf{w}^\top(\boldsymbol{\mu} - r_f \mathbf{1}) - \tfrac{1}{2}\mathbf{w}^\top \boldsymbol{\Sigma} \mathbf{w} \quad \text{s.t.} \quad \mathbf{1}^\top \mathbf{w} = f, \; w_i [...]
$$
with $f \in \{0.5, 1, 2, 3\}$ for Half / Full / Double / Triple Kelly. Long-only clips $f$ to $[0,1]$; unconstrained allows shorts with leverage capped at $|f|\le 3$. The QP is solved via `quadpro[...]

# Results

## Monte Carlo validation under known parameters

The four-horizon Monte Carlo isolates Kelly geometry from estimation noise. At 10,000 trades — roughly 40 years of daily trading — the divergence between mean and median terminal wealth across[...]

```{r mc-table1}
#| label: tbl-mc
#| tbl-cap: "Monte Carlo terminal wealth at 10,000 trades (1,000 paths)."

library(knitr)
mc_tbl <- data.frame(
  Strategy   = c("Half Kelly", "Full Kelly", "Double Kelly", "Triple Kelly"),
  Mean       = c(7.03, 31.47, 183.80, 518.50),
  Median     = c(4.54, 6.89, 1.67, 0.01),
  SD         = c(9, 94, 1803, 11867),
  `P(Loss)`  = c("4.0%", "12.3%", "44.5%", "79.0%"),
  `P(W>2)`   = c("83.6%", "76.3%", "48.3%", "17.5%"),
  `P(W>10)`  = c("20.0%", "41.7%", "29.5%", "10.8%"),
  check.names = FALSE
)
kable(mc_tbl, booktabs = TRUE, align = "lrrrrrr")
```

Under log-normal wealth, $\mathbb{E}[\log W] < \log \mathbb{E}[W]$ by Jensen's inequality and the gap widens with $f^2\sigma^2$. Aggressive multiples buy a larger mean by accepting a much smaller [...]

## Single-equity replication: Banca Intesa, 2007–2018

The Banca Intesa backtest (ISP.MI, daily, 2007–2018) reproduces CC20's published single-asset ranking: Full Kelly delivers the highest terminal wealth (1.193, CAGR +1.53%) at $-23.4\%$ max drawd[...]

## In-sample EuroStoxx 50, 2000–2018

The in-sample multi-asset Kelly portfolio loads 62.3% on ITX.MC and 37.7% on VIV.PA — a two-asset corner solution typical of unconstrained Kelly under sample-mean inputs — and posts 19.07% an[...]

## Out-of-sample EuroStoxx 50: long-only

Once we impose the long-only budget cap $\sum w_i = f \le 1$, the constraint binds for any $f \ge 1$ and Full, Double, and Triple Kelly *collapse onto an identical portfolio* — same QP solution[...]

```{r oos-long-table}
#| label: tbl-oos
#| tbl-cap: "Long-only OOS performance, EuroStoxx 50, 2005-2018. Full / Double / Triple Kelly are identical: the budget constraint binds for $f \\ge 1$."

oos_tbl <- data.frame(
  Strategy        = c("Half Kelly", "Full Kelly", "Double Kelly", "Triple Kelly", "MinVar"),
  Sharpe          = c(0.556, 0.621, 0.621, 0.621, 0.480),
  `Max DD`        = c(0.352, 0.639, 0.639, 0.639, 0.580),
  `Ann Return`    = c("7.87%", "16.87%", "16.87%", "16.87%", "8.55%"),
  `Ann Vol`       = c("14.15%", "27.17%", "27.17%", "27.17%", "17.79%"),
  Sortino         = c(0.835, 0.925, 0.925, 0.925, 0.685),
  `CVaR (0.95)`   = c(0.0195, 0.0380, 0.0380, 0.0380, 0.0261),
  check.names = FALSE
)
kable(oos_tbl, booktabs = TRUE, align = "lrrrrrr")
```

A naive reading endorses CC20: aggressive Kelly variants beat MinVar on Sharpe, return, and Sortino, with the collapsed-aggressive group reaching ~6× by 2018 against ~2.5× for Half Kelly and Mi[...]

## Statistical inference: HAC pairwise tests

Newey-West HAC tests (5 lags) on daily return differences $d_t = r_t^A - r_t^B$ deflate the apparent gaps. Only Half-vs-Full reaches conventional significance ($p=0.014$), and even there the mean[...]

```{r hac-table}
#| label: tbl-hac
#| tbl-cap: "HAC (Newey-West, 5 lags) tests on daily return differences, long-only OOS."

hac_tbl <- data.frame(
  Pair         = c("Half Kelly vs Full Kelly", "Half Kelly vs MinVar", "Full Kelly vs MinVar"),
  `Mean Diff`  = c("-0.0004", "0.0000", "0.0003"),
  `HAC SE`     = c("0.0001", "0.0002", "0.0002"),
  `t-stat`     = c("-2.4646", "-0.1785", "1.4736"),
  `p-value`    = c("0.0138", "0.8584", "0.1407"),
  check.names = FALSE
)
kable(hac_tbl, booktabs = TRUE, align = "lrrrr")
```

The Half-vs-Full significance is mechanically driven by the constraint collapse: Full Kelly's "outperformance" is the difference between an $f = 0.5$ and an $f = 1$ corner solution under the same[...]

## Out-of-sample EuroStoxx 50: unconstrained

Lifting the budget cap separates the four Kelly variants but degrades all of them. Sharpe ratios fall to 0.10–0.18; max drawdowns climb to 72% (Half), 85% (Full), 96% (Double), 97% (Triple); an[...]

## Generalization: Dow Jones Industrial Average

On the DJI single-index backtest (2010–2018), Buy-and-Hold posts a $-20\%$ max drawdown; Half and Full Kelly Long-Only also hold $-20\%$; the unconstrained Half through Triple Kelly variants ru[...]

# Discussion and Conclusion

Three findings emerge that CC20 do not report. *(F1)* The Monte Carlo establishes that Kelly's mean-vs-median divergence is structural, not estimation-driven: aggressive multiples destroy median [...]

Backtest comparisons are inferential exercises, and the absence of formal hypothesis tests in much of the practitioner literature lets cumulative-wealth panels do work they cannot honestly suppor[...]

# References

Breiman, L. (1961). Optimal gambling systems for favorable games. *Proceedings of the Fourth Berkeley Symposium on Mathematical Statistics and Probability*, 1, 65–78.

Carta, A., & Conversano, C. (2020). Practical implementation of the Kelly Criterion. *Frontiers in Applied Mathematics and Statistics*, 6, 577050.

Cover, T. M. (1991). Universal portfolios. *Mathematical Finance*, 1(1), 1–29.

Kelly, J. L. (1956). A new interpretation of information rate. *Bell System Technical Journal*, 35(4), 917–926.

MacLean, L. C., Thorp, E. O., & Ziemba, W. T. (Eds.). (2011). *The Kelly Capital Growth Investment Criterion*. World Scientific.

Newey, W. K., & West, K. D. (1987). A simple, positive semi-definite, heteroskedasticity and autocorrelation consistent covariance matrix. *Econometrica*, 55(3), 703–708.

Thorp, E. O. (2011). The Kelly criterion in blackjack, sports betting, and the stock market. In *The Kelly Capital Growth Investment Criterion* (pp. 789–832). World Scientific.
