---
title: "Appendix - Supporting Figures and Source Code"
author: "Syed Bashir Hydari et al."
date: today
format:
  pdf:
    toc: true
    toc-depth: 1
    number-sections: true
    fig-width: 7
    fig-height: 4.2
    fig-pos: "H"
    geometry:
      - margin=1in
    include-in-header:
      text: |
        \usepackage{float}
        \usepackage{booktabs}
        \usepackage{caption}
        \captionsetup[figure]{font=small,skip=2pt}
        \captionsetup[table]{font=small,skip=2pt}
        \setlength{\parskip}{4pt}
        \renewcommand{\thefigure}{A\arabic{figure}}
        \renewcommand{\thetable}{A\arabic{table}}
execute:
  warning: false
  message: false
  echo: true
  eval: true
  cache: true
editor: source
---

\newpage

# Overview

This appendix presents seven supporting figures referenced in the main report alongside the executable source code that produces each one. Section 2 sets up the global R environment. Sections 3–[...]

# Setup

The block below loads the package stack used throughout the appendix and applies a namespace fix that forces `dplyr`'s `filter`, `lag`, and `select` to win over the masking versions exported by `x[...]

```{r setup}
suppressPackageStartupMessages({
  library(tidyverse)
  library(scales)
  library(patchwork)
  library(knitr)
  library(quantmod)
  library(tidyquant)
  library(xts)
  library(zoo)
  library(quadprog)
  library(Matrix)
  library(PerformanceAnalytics)
  library(portfolioBacktest)
  library(sandwich)
  library(lmtest)
})

# Force dplyr verbs to win over xts/stats/PerformanceAnalytics masks.
filter <- dplyr::filter
lag    <- dplyr::lag
select <- dplyr::select

options(scipen = 999)
options(knitr.kable.NA = "")
theme_set(theme_minimal(base_size = 11))

GLOBAL_SEED <- 123
set.seed(GLOBAL_SEED)
```

\newpage

# Figure A1 — Monte Carlo Wealth Trajectories (§3.1)

Sample wealth trajectories under each Kelly multiple over 1,000 simulated trades. The visual signature of the volatility-pumping geometry discussed in §3.1 of the main report: Half Kelly's paths [...]

```{r mc-parameters}
MU_ANNUAL    <- 0.12
SIGMA_ANNUAL <- 0.40
RF_ANNUAL    <- 0.01
TRADING_DAYS <- 252

MU_DAILY    <- MU_ANNUAL / TRADING_DAYS
SIGMA_DAILY <- SIGMA_ANNUAL / sqrt(TRADING_DAYS)
RF_DAILY    <- RF_ANNUAL / TRADING_DAYS

F_FULL_KELLY   <- (MU_DAILY - RF_DAILY) / (SIGMA_DAILY^2)
F_HALF_KELLY   <- 0.5 * F_FULL_KELLY
F_DOUBLE_KELLY <- 2.0 * F_FULL_KELLY
F_TRIPLE_KELLY <- 3.0 * F_FULL_KELLY
```

\newpage

# Figure A2 — Final Wealth Distributions Across Horizons (§3.1)

Density estimates of terminal wealth at 100, 1,000, and 10,000 simulated trades, with dashed lines marking medians. This figure operationalizes the mean-vs-median story: at 10,000 trades, Triple [...]

\newpage

# Figure A3 — Banca Intesa Cumulative Wealth (§3.2)

Cumulative wealth paths for Buy-and-Hold and three Kelly multiples on Banca Intesa (ISP.MI, 2007–2018). The Kelly fraction is estimated once from the full sample as $f^* = (\hat{\mu} - r_f)/\ha[...]

\newpage

# Figure A4 — Mean-Variance Efficient Frontier, In-Sample EuroStoxx 50 (§3.3)

The mean-variance plane for the in-sample EuroStoxx 50 panel (2000–2018, monthly returns), with the four optimal portfolios overlaid: equal-weight, minimum-variance, tangent (max-Sharpe), and K[...]

\newpage

# Figure A5 — Long-Only OOS Cumulative Returns, EuroStoxx 50 (§3.4)

Out-of-sample cumulative wealth for the five long-only strategies on EuroStoxx 50 (2005–2018), produced by a rolling-window backtest with 756-day window and monthly re-optimization. The visual [...]

\newpage

# Figure A6 — HAC Pairwise Return Differences (§3.5)

This is the project's central inferential extension. For each pair of strategies $(A, B)$ we form the daily return-difference series $d_t = r_t^A - r_t^B$, regress on a constant, and plot the mea[...]

\newpage

# Figure A7 — Unconstrained Allocation Stack, Half Kelly (§3.6)

Asset weights through time for the Half Kelly variant under the unconstrained QP (shorts permitted, leverage capped at $|f| \le 3$). The plot is the visual evidence for the over-leverage patholog[...]
